\part{Principles of game development}
\chapter{Beta testing}
\section{In General}

Beta testing is perhaps the most important phase of any video game's development. Whether you've spent a month or a few years on your adventure, if you don't have it beta tested properly, all that work could be ruined by one simple little gameplay killer. If you understand the importance of beta testing, then read this tutorial carefully and take it to heart. \emph{If you don't understand the importance of beta testing, then read this tutorial carefully and take it to heart.} I cannot stress enough how important the beta testing phase is.
\par When should beta testing begin? The short answer is after alpha testing is finished. When is alpha testing finished? That's easy! That's when you think your game is ready for release. Your game is the best you can get it and you're sure there is no more you can do and you feel happy about releasing it. Now it's time for beta testing. This seems to be a difficult concept for new Level Builders (and a few seasoned Level Builders) to get their heads around. Trust me, your game does not go for beta testing until you are convinced it's finished. If it's not finished, you're still in the alpha testing stages and if you employ your beta testers now they will be wasting their time and your game will not be beta tested properly.
\par When a game moves from alpha testing to beta testing, the only changes you should make would be to fix bugs, gameplay errors and textural errors. You should never build new areas or new gameplay or new puzzles during or after beta testing. Again I stress, do not begin beta testing until your game is FINISHED. Beta testing is purely to iron out problems in finished games. If you ignore this advice the chances are you will release a buggy game. Beta testers take their job seriously and if games are released with bugs it reflects on them and the job they did. If those bugs are down to the Level Builder changing things during or after beta testing and the game is released with bugs you will never get beta testers to help you again. Beta testers will invest days and even weeks of their personal time to help you with your game. Respect that and don't let them down.

\section{Alpha Testing}
Don't cut corners and skip this step. Yes, I know, you're excited about your game and you want to show it off and that's marvellous and wonderful and admirable and quite understandable, but your beta testers are not there to do your job for you. If you upload an appallingly buggy game that hasn't been properly alpha tested, don't be surprised if your beta testers don't do a good job and don't volunteer later when you actually need beta testers. Alpha testing is your job, not your beta testers, and it's extremely important. I consider my alpha testing to be complete when I have a package I think is ready for release. In other words, I can find nothing to change or fix and in my heart I know my game is the best I can make it and I feel it's ready for release. Now I begin beta testing.
\par During alpha testing, it is easy to move Lara around and forget to put her back, place temporary triggers to open a door and then forget to remove them and a few other little things, so after I've packaged up my game for beta testing guess what? Yes, I then test the beta package from start to finish using only the files I will be uploading for my beta testers. This is my final Alpha test. If this goes well and all the level jumps work properly and I successfully hit a finish trigger, only then is the game uploaded and the download link sent out to the beta testers.

\cite{NGLE_manual_hu}